Notes of Basic Topolgy

### 2016-12-23 Basic Topology Notes/sections/ch6-Triangulation.tex
[major] Introducing simplicial homology groups.
[major] Examples of homology groups, and introduce chain complex.

\begin{defi}[Triangulation and Triangulable]
\nomenclature{Triangulation and Triangulable}{\nomrefpage.}
    A triangulation on a topological space $X$ consists of a
    simplicial complex $K$, and a homeomorphism $h$:
    $$ h:|K|\to X$$
    A space is called triangulable if such simplicial complex $|K|$
    and homeomorphism $h$ exists.
\end{defi}
\begin{fact}
    If $X$ is triangulable, then $X$ is compact and can be made into a
    metric space. (Notice that the simplicial complex is a finite
    collection.)
\end{fact}
\begin{remark}
    Trigulation is in general not unique. For example:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/rmk2}
    \end{figure}
\end{remark}
\begin{ex}
    Here is a triangulation on the torus:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/p5.pdf}
    \end{figure}
    Notice that the following diagram is not a triangulation of torus:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/ch6-scanned-notes-1/p52.pdf}
    \end{figure}
\end{ex}
\begin{lemma}
    Let $K$ be a simplicial complex in $\R^{n}$, then
    \begin{enumerate}
        \item $|K|$ is compact.
        \item Each point of $|K|$ lies in the interior of exactly one
            simplex of $K$.
        \item If we take the simplexes of $K$ seperately, and give
            their union the identification topology, we obtain $|K|$.
            Thus we have two equivalent ways to view the topology on
            $K$.
        \item If $|K|$ is connected, then $|K|$ is also
            path-connected.
    \end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate}
    \item Obviously.
    \item Since $K$ is connected by simplexes, a point $p$ must
        obviously be contained in some simplex. Also, since on each
        simplex $p$ must be on the interior of it, or interior of its
        components. We only need to prove that the containing simplex
        is unique. Suppose $p$ is the interior of simplexes $A$ and
        $B$. Notice that their intersection must be their faces. The
        only face of $A$ or $B$ containing interior points is $A$ or
        $B$ itself. Hence $A=B$.
    \item By definition of identification topology, we can see that in
        the identified space, a subset $C\subset |K|$, $C$ is closed
        $\Leftrightarrow$ $C\cap |A|$ is closed in $|A|$ for all
        simplexes $A\in K$. The rest of the proof is obvious.
    \item We need only to show that $|K|$ is locally path-connected.
        But any point $p\in K$, $p\in$ interior of some $A$ where
        $A\in K$. That is, we can find a neighbourhood $U$ of $p$, and
        $U$ is contained in $A$ ($U\subset A$). (More specifically, we
        can find $\varepsilon$ such that 
        $B_{\varepsilon}(p)\cup |K| = B_\varepsilon(p)\cup |A|$)
        But a simplex $A$ is obviously locally path-connected. Hence
        $U$ (or $B_\varepsilon(p)$) is path connected, hence $|K|$ is
        path-connected.
\end{enumerate}
\end{proof}

@@

\subsection{6.4 The edge group of a complex}
\label{sec:edge-group}

@@

\begin{ex}
    Here is the maximal tree of the triangulation of two-torus:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\linewidth]{pics/ch6-notes-3/ex1-maximal-tree.png}
    \end{figure}
\end{ex}

@@

\begin{ex}
    \label{ex:torus-GLKgroup}
    Again, for torus:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\linewidth]{pics/ch6-notes-3/ex3-L-in-triangle.png}
    \end{figure}
\end{ex}
\begin{ex}
    \label{ex:ex4-L-in-two-torus}
    And for another:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\linewidth]{pics/ch6-notes-3/ex4-L-in-two-torus.png}
    \end{figure}
    Note that this does not gives a triangulation of two-torus! 
    The reason will be explained later in
    example~\ref{ex:pi1-two-torus}.
\end{ex}
More generally, we have
\begin{thm}
    $G(K,L)\cong E(K,v)$, as a group isomorphism.
\end{thm}
\begin{proof}
    The proof in page 135 to 136 of \cite{book} constructs the
    following homomorphisms:

>>>>

\begin{defi}[Triangulation and Triangulable]
\nomenclature{Triangulation and Triangulable}{\nomrefpage.}
    A triangulation on a topological space $X$ consists of a
    simplicial complex $K$, and a homeomorphism $h$:
    $$ h:|K|\to X$$
    A space is called triangulable if such simplicial complex $|K|$
    and homeomorphism $h$ exists.
\end{defi}
\begin{fact}
    If $X$ is triangulable, then $X$ is compact and can be made into a
    metric space. (Notice that the simplicial complex is a finite
    collection.)
\end{fact}
\begin{remark}
    Trigulation is in general not unique. For example:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/rmk2}
    \end{figure}
\end{remark}
\begin{ex}
    \label{ex:triangu-torus}
    Here is a triangulation on the torus:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch6-scanned-notes-1/p5.pdf}
    \end{figure}
    Notice that the following diagram is not a triangulation of torus:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/ch6-scanned-notes-1/p52.pdf}
    \end{figure}
\end{ex}
\begin{lemma}
    Let $K$ be a simplicial complex in $\R^{n}$, then
    \begin{enumerate}
        \item $|K|$ is compact.
        \item Each point of $|K|$ lies in the interior of exactly one
            simplex of $K$.
        \item If we take the simplexes of $K$ seperately, and give
            their union the identification topology, we obtain $|K|$.
            Thus we have two equivalent ways to view the topology on
            $K$.
        \item If $|K|$ is connected, then $|K|$ is also
            path-connected.
    \end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate}
    \item Obviously.
    \item Since $K$ is connected by simplexes, a point $p$ must
        obviously be contained in some simplex. Also, since on each
        simplex $p$ must be on the interior of it, or interior of its
        components. We only need to prove that the containing simplex
        is unique. Suppose $p$ is the interior of simplexes $A$ and
        $B$. Notice that their intersection must be their faces. The
        only face of $A$ or $B$ containing interior points is $A$ or
        $B$ itself. Hence $A=B$.
    \item By definition of identification topology, we can see that in
        the identified space, a subset $C\subset |K|$, $C$ is closed
        $\Leftrightarrow$ $C\cap |A|$ is closed in $|A|$ for all
        simplexes $A\in K$. The rest of the proof is obvious.
    \item We need only to show that $|K|$ is locally path-connected.
        But any point $p\in K$, $p\in$ interior of some $A$ where
        $A\in K$. That is, we can find a neighbourhood $U$ of $p$, and
        $U$ is contained in $A$ ($U\subset A$). (More specifically, we
        can find $\varepsilon$ such that 
        $B_{\varepsilon}(p)\cup |K| = B_\varepsilon(p)\cup |A|$)
        But a simplex $A$ is obviously locally path-connected. Hence
        $U$ (or $B_\varepsilon(p)$) is path connected, hence $|K|$ is
        path-connected.
\end{enumerate}
\end{proof}

@@

\subsection{Section 6.4 The edge group of a complex}
\label{sec:edge-group}

@@

\begin{ex}
    Here is the maximal tree of a triangulation:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch6-notes-3/ex1-maximal-tree.png}
    \end{figure}
\end{ex}

@@

\begin{ex}
    \label{ex:torus-GLKgroup}
    Again, for circle:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\linewidth]{pics/ch6-notes-3/ex3-L-in-triangle.png}
    \end{figure}
\end{ex}
\begin{ex}
    \label{ex:ex4-L-in-two-torus}
    And for another:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\linewidth]{pics/ch6-notes-3/ex4-L-in-two-torus.png}
    \end{figure}
    Note that this does not gives a triangulation of two-torus! 
    The reason will be explained later in
    example~\ref{ex:pi1-two-torus}.
\end{ex}
More generally, we have
\begin{thm}
    $G(K,L)\cong E(K,v)$, as a group isomorphism.
\end{thm}
\begin{proof}
    The proof in page 135 to 136 of \cite{book} constructs the
    following homomorphisms:

### 2016-12-23 Basic Topology Notes/sections/ch8-Simplicial-Homology.tex
[major] Examples of homology groups, and introduce chain complex.

% todo torus-p1
\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/ch8/torus-p1.png}
\end{figure}

@@

I only mention two theorems for the present:

@@

>>>>

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\linewidth]{pics/ch8/torus-p1.png}
\end{figure}

@@

I will mention two theorems first:

@@

The two theorems indicates again that those homology groups are
independent of the choice of triangulation, and they should be
topological invariants.

Then I will turn attention to specific examples.

\begin{key}
    \label{key:2cycle-orientation}
    Observe that every cycle is actually a closed object. It encircles
    or encloses an aresa. A closed $1$-cycle is actually a loop. A
    closed $2$-cycle will always enclose an area, so it is actually a
    sphere. A closed $2$-cycle will give an orientation on its
    surface. For example, suppose that we have a closed $2$-cycle
    $\lambda_i\sigma_i$, and assume that there is no repeated
    $2$-simple in this collection. For a given $2$-simplex $\sigma_1$
    in it, there will have to be $3$ simplexes surrounding it to
    cancal out $\sigma_1$'s edges. These $4$ $2$-simplexes are said to
    have compatible orientation. If one draw them on a paper, one will
    see that they are of the same orientation. Continue this argument
    to include all simplexes and one will see this $2$-cycle actually
    gives an orientation on its surface.
\end{key}
\begin{ex}\textbf{Torus}

    Consider the triangulation of torus as in
    exmaple~\ref{ex:triangu-torus}.
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch8/torus-tri.png}
    \end{figure}

    The zeroth homology group $H_0(K)$ marks the connected components,
    hence it is $H_0(K)=\Z$. The first homology group $H_1$ is about
    those closed loops. Almost all loops are boundaries. But there are
    two (the green one and the blue one), each of which cannot be a
    boundary. For example, I managed to make the orientation such that
    the green one is not cancelled, but this leaves a part of the blue
    one also un-cancelled. 
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch8/torus-tri-1.png}
    \end{figure}

    This two loops generates the $H_1(K)=\Z\cross\Z$, and are also the
    nontrivial objects in its fundamental group (a coindence,
    $\pi_1(T^2)=H_1(T^2)$).

    The $2$-cycles are easy to find. Because our space torus has no
    boundary, we have to add more and more $2$-simplexes until the
    whole space is covered with a $2$-chain. Then in this case, there
    is only one generator for $Z_2(K)$. Also, there is not
    $3$-simplex, so the $B_2=0$. Hence $H_2(K)=Z_2(K)=\Z$.
\end{ex}

\begin{ex} \textbf{Klein bottle $T^2$}

    As for the Klein bottle, I have had a hard time getting and
    confirming that the following is a triangulation:
    \footnote{
        I found two articles talking about minimal triangulations
        of Klein bottle. They are:
        \href{https://arxiv.org/abs/math/0407008}{link1}
        and
        \href{http://www.sciencedirect.com/science/article/pii/S0095895697999998}{link2}.
        The pdf files is also included.
    }
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/ch8/klein-tri.png}
    \end{figure}

    The $H_0(K)$ is clearly $\Z$.

    The $1$-cycles on this triangulation are hard to analyse. So I
    will use the fact that its $\pi_1=\braket{a,b|ba=a^{-1}b}$.
    The commutator is generated by $aba^{-1}b^{-1}=a^2$ and
    $bab^{-1}a^{-1}=a^{-2}$. So in the abelianization,
    $[a]=[a^{2n+1}]$, $[a^{2n}]=[1]$. So $[a]$ generates $\Z_2$. The
    $[b]$ is not affected, so $[b]$ generates $\Z$. It's easy to
    confirm that $[ab]=[ba]$, i.e. the abelianization group is
    abelian. All in all, $H_1(K)=\Z\cross\Z_2$.

    The $H_2(K)=0$, since there is no way to orient the Klein in a
    reasonable way (see key point~\ref{key:2cycle-orientation}).
\end{ex}

\begin{ex}\textbf{Cone Space}

    Now the book considers a complex $K$ which is a cone, i.e. $K$ is
    isomorphic to a complex of the form $CL$ where the dimentions of
    $L$ is $(\dim{K}-1)$.

    It is easy to see that $H_0(K)=\Z$, since cone is always
    connected.
    It is shown in page 181 of book\cite{book} that $H_n(K)=0$, for $n\geq
    1$. This is reasonable because cones are actually contractible, so
    there is not much interesting information about higher dimensional
    spheres in cones to be recorded. The book shows this by
    constructing a map $d:C_q(K)\to C_{q+1}(K)$, such that
    \begin{equation}
        \partial d(\sigma) = \sigma-d\partial(\sigma)
    \end{equation}
    It's easy to see that this means every closed cycle is a boundary.
\end{ex}

\begin{ex}\textbf{$(n+1)$-simplexes and $\sum^n\cong S^n$}

    Let $\Delta^{n+1}$ denotes a $(n+1)$-simplex. In this case, we
    consider it as a simplicial complex with all of its faces. Let
    $\sum^n$ denotes the simplexes which lies in the boundary of this
    $(n+1)$-simplex $\Delta^{n+1}$. Notice this is equivalent to say
    that $\sum^n$ is the all remaing simplex obtained by removing the
    interior points of $\Delta^{n+1}$. For example, $\sum^1$ is the
    3 edges and 3 vertices of the triangle. We see that $\sum^n$ is
    isomorphic to $S^n$.

    Every $\Delta^{n+1}$ can be thought of as a cone (except when
    $n=-1$, which is too trivial to be discussed here). So the
    homology classes of $\Delta^{n+1}$ are $H_0(\Delta^{n+1})=\Z$,
    $H_q(\Delta^{n+1})=0$, $q\geq 1$.

    The homology classes of $\sum^n$ can be calculated using
    $\Delta^{n+1}$, because they differ by only a single
    ${n+1}$-simplex, then most of their chains are the same.

    When $n=0$, the case is trivial and $H_0(\sum^0)=\Z\oplus\Z$,
    $H_q(\sum^0)=0$ for $q\geq 0$.

    When $n\geq 1$, then since $H_q$ does not involve simplexes of
    dimension higher than $q+1$, we have
    $H_q(\sum^n)=H_q(\Delta^{n+1})$, for $0\leq q\leq n-1$. Hence
    $H_0(\sum^n)=\Z$. $H_q(\sum^n)=0$, for $1\leq q \leq n-1$. 
    By using $H_n(\Delta^{n+1})=0$, it is easy to get
    $H_n(\sum^n)=\Z$. (book\cite{book} page 182). The rest of the Homology groups
    will of course be $0$.

    It is interesting to check that $H_n(\sum^n)=\Z$, not $0$ for a
    one point space. If we are able to show that Homology group is a
    topological invariant, then we know now Homology groups can be
    used to distinguish spheres with a point.
\end{ex}

\subsection{Section 8.4 Simplicial maps}
\label{sec:Simplicial maps}

Now we extend a simplicial map $K\to L$ to a homomorphism between
homology groups. This is easily done by first constructing
$s:C_q(K)\to C_q(L)$, and then confirming $s\partial=\partial s$, or
just $[s,\partial]=0$. This is a very general procedure and the
condition will be frequently seen in Homological algebra or
Cohomological algebra. In face, let me first introduces Homological
concepts first.

The collection of groups and homomorphisms:
$$\begin{tikzcd}[]
    \cdots \ar[r,"\partial"] & C_q(K) \ar[r,"\partial"] &
    C_{q-1}(K) \ar[r,"\partial"] & \cdots \ar[r,"\partial"] &
    C_0(K) \ar[r,"\partial"] & 0
\end{tikzcd}$$
will be called a \nomen{chain complex} of $K$, written $C(K)$. 
A \nomen{chain map} $\phi:C(K)\to C(L)$ is a set of group
homomorphisms $\phi:C_q(K)\to C_q(L)$. Such a map will induce a
homomorphism of Homology groups when $\phi\partial=\partial\phi$
(easily proved). This induced homomorphism between Homology groups is
dentoed $\phi_*$.

In our specific case where $s$ are simplicial maps $K\to L$, it is
easy to extend it to $s:C_q(K)\to C_q(L)$, since $s$ has already been
required to map simplexes to simplexes. We only need to pay attention
when $\sigma$ a generator (i.e. a $q$-simplex) of $C_q(K)$, is mapped
to a simplex of inferior quaility, i.e. of lower dimension $<q$. Then
it is nature to define $s(\sigma)=0$ in this case. It is verifed (page
185 of \cite{book}) that this map is well defined, is a homomorphism,
and has the property $s\partial=\partial s$. So $s$ induces a
homomorphism $s_*:H_q(K)\to H_q(L)$.
